---
title: "Homogeneity and normality"
author: "Robby"
date: "7/31/2020"
output: 
    html_document:
        keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(FSA)
library(car)
library(nortest)

```


### **Reading and preparing data**
```{r}
df <- read.delim("Anggrek Tanah (Tinggi).txt")
kable(df, format = "html")
dependent.var <- df[,2] #Column variabel dependent
independent.var <- df[,1]#Column variabel independent
satuan <- "(cm)"
```

### **Data descriptive summary**
```{r message=FALSE, warning=FALSE}
attach(df)
```

```{r}
df_summary <- data.frame(Summarize(Tinggi.Tanaman ~ Media, digits = 1))
kable(df_summary, format = "pipe")
```

### **Homogeneity test**

#### Levene's test for homogeneity of variance.
*if p > 0.05 assumption of homogeneity fulfilled*
```{r}
levene <- suppressWarnings(leveneTest(Tinggi.Tanaman ~ Media, data = df))
print(levene)
```

### **Normality test**
*if p > 0.05 assumption of normal distribution fulfilled*

#### Shapiro-Wilk Normality Test
```{r}
normalitas <- shapiro.test(Tinggi.Tanaman)
print(normalitas)
```

#### Kolmogorov-Smirnov Tests
```{r}
uji_ks <- suppressWarnings(ks.test(Tinggi.Tanaman,"pnorm",mean = mean(Tinggi.Tanaman), 
                            sd = sd(Tinggi.Tanaman)))
print(uji_ks)
```

#### Lilliefors (Kolmogorov-Smirnov) test for normality
Although the test statistic obtained from lillie.test(x) is the same as that obtained from ks.test(x, "pnorm", mean(x), sd(x)), it is not correct to use the p-value from the latter for the composite hypothesis of normality (mean and variance unknown), since the distribution of the test statistic is different when the parameters are estimated.

```{r}
lillie <- lillie.test(Tinggi.Tanaman)
print(lillie)
```

### **Q-Q Plot**
A Q-Q plot is a scatterplot created by plotting two sets of quantiles against one another. If both sets of quantiles came from the same distribution, we should see the points forming a line that’s roughly straight.
```{r}
kol2 <- colnames(df)[2]
QQplot <- qqPlot(Tinggi.Tanaman, id = F, ylab = kol2, main = "Quantile - Quantile Plot")
```

### **Box-and-whisker plot**
```{r fig.align='left'}
box_whisker <- boxplot(Tinggi.Tanaman ~ Media, 
                        xlab = paste(kol2,satuan), ylab = "", col = "orange", border = 
                        "brown", horizontal = TRUE, las = 1, par(mar = c(8,13,1,1)))
```


