---
title: "Diversity Measures (beta and abundance)"
subtitle: "PT ANJ Jan - Jun 2020"
author: "Robby B"
date: "8/7/2020"
output: 
  html_document:
    toc: TRUE
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load packages**

```{r message=FALSE, warning=FALSE}
library(tabula)
library(magrittr)
library(khroma)
library(ggplot2)
library(openxlsx)
library(RFLPtools)
library(knitr)
```

**Reading and preparing data**

```{r}
dat <- "data_jml_spesies_ANJ.xlsx"
df <- read.xlsx(paste0("C:/Users/rbbutar/Documents/R/Konservasi_ANJ/indeks_kehati/", dat), 
                sheet = "grafik_beta", cols = 1:8)
rownames(df) <- df[,1]
df <- df[,-1]
df <- t(df)

#Dropping columns sum 0
df <- as.data.frame(df)
df <- df[,!sapply(df, function(x) sum(x))==0]
kable(df, format = "pipe")
```


## A. Beta -diversity

Beta diversity is a comparison of diversity between ecosystems, usually measured as the amount of species change between the ecosystem.

### 1. Similarity

Jaccard, Morisita-Horn and Sorenson indices provide a scale of similarity from 0-1 where 1 is perfect similarity and 0 is no similarity. The Brainerd-Robinson index is scaled between 0 and 200.

### a. Brainerd-Robinson (similarity between assemblages)

- **Quantitative** similarity measures (between samples)
- Brainerd-Robinson quantitative index. This is a city-block metric of similarity between pairs of samples/cases.

```{r}
brainerd <- df %>% as_count %>% similarity(method = "brainerd")
brainerddata <-sim2dist(brainerd@.Data, maxSim = 200)
plot(
    hclust(brainerddata),
    hang = -1,
    main = "Sites clustered by Brainerd similarity",
    xlab = "Quantitative Index",
    axes = FALSE, ylab = ""
    )
```

```{r}
brainerd %>% plot_spot() + khroma::scale_colour_YlOrBr() + 
            theme(axis.text = element_text(size = 12, face = "bold"))
```

### b. Jaccard index (similarity between assemblages)

- **Qualitative** similarity measures (between samples)
-  This analysis includes presence/absence standardization using `decostand`

```{r}
jaccard <- df %>% as_count %>% similarity(method = "jaccard")
jaccarddata <- sim2dist(jaccard@.Data, maxSim = 1)
plot(
    hclust(jaccarddata),
    hang = -1,
    main = "Sites clustered by Jaccard similarity",
    xlab = "includes presence/absence standardization using `decostand`",
    sub = "Qualitative index",
    axes = FALSE, ylab = ""
    )
```

```{r }
jaccard %>% plot_spot() + khroma::scale_colour_BuRd() + 
    theme(axis.text = element_text(size = 12, face = "bold"))
```

```{r, include=FALSE}
### c. Binomial co-occurrence (similarity between types)

#Binomial co-occurrence assessment. This assesses the degree of co-occurrence between taxa/types within a dataset. The strongest associations are shown by large positive numbers, the strongest segregations by large negative numbers.

#- Similarity between pairs of taxa/types can be measured by assessing the degree of co-occurrence (binomial co-occurrence).
#- The Binomial co-occurrence assessment approximates a Z-score.


#sim.types <- df %>% as_count %>% similarity(method = "binomial") %>%
           # plot_spot() + khroma::scale_colour_PRGn() +
           # theme(axis.text = element_text(size = 5, face = "bold"))
#print(sim.types)
```

## B. Abundance Model

Ranks vs abundance plot can be used for abundance models

```{r}
abund.model <- df %>% as_count() %>% 
            plot_rank(log = "xy", facet = FALSE) +
            ggplot2::theme_bw() + khroma::scale_color_discreterainbow()
print(abund.model)
```


