---
title: "Species_modelling"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("sp")
library("raster")
library("maptools")
library("rgdal")
library("dismo")
library("sf")
library("terra")
library("geodata")
``` 


```{r}
#bioclim.data <- getData(name = "worldclim",
                        #var = "bio",
                        #res = 5,
                        #path = "~/My R/Konservasi_ANJ/species_modelling")

bioclim.data <- worldclim_global("bio", 5, "~/My R/Konservasi_ANJ/species_modelling/output/5_res/worldclimbaru")
bioclim.data <- brick(bioclim.data)

```

```{r}
obs.data <- st_read("~/My R/Konservasi_ANJ/species_modelling/white_tailed_eagle.shp")
names(obs.data)[22]<- "latitude"
names(obs.data)[23] <- "longitude"

```

To make species distribution modeling more streamlined, it is useful to have an idea of how widely our species is geographically distributed. We are going to find general latitudinal and longitudinal boundaries and store this information for later use:
```{r}
#Determine geographic extent of our data
max.lat <- ceiling(max(obs.data$latitude))
min.lat <- floor(min(obs.data$latitude))
max.lon <- ceiling(max(obs.data$longitude))
min.lon <- floor(min(obs.data$longitude))
geographic.extent <- extent(x = c(min.lon, max.lon, 50, 61))

```

```{r, fig.dim= c(6,8)}
# Load the data to use for our base map
great_britain <- st_read("~/My R/Konservasi_ANJ/species_modelling/greatbritain.shp")

data("wrld_simpl")

great_britain_prj <- st_transform(great_britain, crs = raster::crs(wrld_simpl, asText = T))

# Plot the base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Add the points for individual observation
points(x = obs.data$longitude, 
       y = obs.data$latitude, 
       col = "olivedrab", 
       pch = 20, 
       cex = 0.75)
# And draw a little box around the graph
box()


```

#Building a model and visualizing results
Now that our occurrence data look OK, we can use the bioclimatic variables to create a model. The first thing we want to do though is limit our consideration to a reasonable geographic area. That is, for our purposes we are not looking to model saguaro habitat suitability globally, but rather to the general southwest region. So we can restrict the biolimatic variable data to the geographic extent of our occurrence data:

```{r}
# Crop bioclim data to geographic extent of saguaro
bioclim.data <- raster::crop(x = bioclim.data, y = geographic.extent)

obs.data <- obs.data[, c("longitude", "latitude")]
obs.data <- as_Spatial(obs.data)

# Build species distribution model
bc.model <- bioclim(x = bioclim.data, p = obs.data@data)

```

```{r}
# Predict presence from model
predict.presence.historical <- dismo::predict(object = bc.model, 
                                   x = bioclim.data, 
                                   ext = geographic.extent)

```

This plot shows the probability of occurrence of saguaros across the map. Note the values are all quite below 1.0; in fact, the maximum probability anywhere on the map is only 0.78, according to the model. However, we are pretty sure that saguaros are found across a pretty broad area of the Sonoran Desert - after all, we have the observations to prove that! If we want our map to better reflect this, we will need to re-run our analyses, but this time include some absence points, where saguaros are known to not occur. The problem is, we only have presence data for saguaros.

```{r, fig.dim= c(6,8)}
# Plot base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Add model probabilities
plot(predict.presence.historical, add = TRUE)

# Redraw those country borders
plot(great_britain_prj, add = TRUE, border = "grey5")

# Add original observations
#points(obs.data$longitude, obs.data$latitude, col = "black", pch = 1, cex = 0.5)
box()

```

#The pseudo-absence point
One common work around for coercing presence-only data for use with presence/absence approaches is to use pseudo-absence, or “background” points. While “pseudo-absence” sounds fancy, it really just means that one randomly samples points from a given geographic area and treats them like locations where the species of interest is absent. A great resource investigating the influence and best practices of pseudo-absence points is a study by Barbet-Massin et al. (2012) (see Additional Resources below for full details).

For our purposes, we are going to create a set of background (aka pseudo-absence) points at random, with as many points as we have observations. We are going to use the bioclim data files for determining spatial resolution of the points, and restrict the sampling area to the general region of the observations of saguaros.

```{r}
# Use the bioclim data files for sampling resolution
bil.files <- list.files(path = "~/My R/Konservasi_ANJ/species_modelling/output/5_res/worldclimbaru/wc2.1_5m", 
                          pattern = "*.bil$", 
                          full.names = TRUE)

# We only need one file, so use the first one in the list of .bil files
mask <- raster(bil.files[1])

# Set the seed for the random-number generator to ensure results are similar
set.seed(20210707)

# Randomly sample points (same number as our observed points)
background <- randomPoints(mask = mask,     # Provides resolution of sampling points
                   n = nrow(obs.data),      # Number of random points
                   ext = geographic.extent, # Spatially restricts sampling
                   extf = 1.25)             # Expands sampling a little bit

```

We can also visualize them on a map, like we did for the observed points:

```{r, fig.dim= c(6,8)}
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95",
     main = "Presence and pseudo-absence points")

# Add the background points
points(background, col = "grey30", pch = 1, cex = 0.75)

# Add the observations
points(x = obs.data$longitude, 
       y = obs.data$latitude, 
       col = "olivedrab", 
       pch = 20, 
       cex = 0.75)

box()

```

Now that we have our pseudo-absence points, we need to take one more step. Getting a more traditional-range-map-looking figure requires post hoc evaluation of the model. To do this evaluation, we are going to build the model using only part of our data (the training data), reserving a portion of the data for evaluation of the model after it is build (the testing data). We are going to reserve 20% of the data for testing, so we use the kfold function in the dismo package to evenly assign each observation to a random group.

```{r}
# Arbitrarily assign group 1 as the testing data group
testing.group <- 1

# Create vector of group memberships
group.presence <- kfold(x = obs.data, k = 5) # kfold is in dismo package
```

Now pause for a minute and take a look at that group.presence vector we just created:

```{r}
head(group.presence)
# Should see even representation in each group
table(group.presence)

```

The output of table shows how many points have been assigned to each of the five groups. In this case, we can see that the points have been evenly distributed, with 20% of the points in group 1, our testing group.

We use the group.presence vector with the observed data to separate our observations into a training data set and a testing data set:

```{r}
# Separate observations into training and testing groups
presence.train <- obs.data[group.presence != testing.group, ]
presence.test <- obs.data[group.presence == testing.group, ]

# Repeat the process for pseudo-absence points
group.background <- kfold(x = background, k = 5)
background.train <- background[group.background != testing.group, ]
background.test <- background[group.background == testing.group, ]

```

#Training and testing the model
Now that we have (1) our pseudo-absence points and (2) separate training and testing data, we can re-build the model, evaluate its performance, and draw a more aesthetically pleasing map. We build the model with the bioclim function as before, but instead of using all the observations in obs.data we only use the training data stored in presence.train:

```{r}
# Build a model using training data
bc.model <- bioclim(x = bioclim.data, p =presence.train@data)

# Predict presence from model (same as previously, but with the update model)
predict.presence.pseudo <- dismo::predict(object = bc.model, 
                                   x = bioclim.data, 
                                   ext = geographic.extent)

```

We now take that model, and evaluate it using the observation data and the pseudo-absence points we reserved for model testing. We then use this test to establish a cutoff of occurrence probability to determine the boundaries of the saguaro range.
```{r}
# Use testing data for model evaluation
bc.eval <- evaluate(p = presence.test,   # The presence testing data
                    a = background.test, # The absence testing data
                    model = bc.model,    # The model we are evaluating
                    x = bioclim.data)    # Climatic variables for use by model

# Determine minimum threshold for "presence"
bc.threshold <- threshold(x = bc.eval, stat = "spec_sens")

```

The threshold function offers a number of means of determining the threshold cutoff through the stat parameter. Here we chose "spec_sens", which sets “the threshold at which the sum of the sensitivity (true positive rate) and specificity (true negative rate) is highest.” For more information, check out the documentation for threshold (?threshold, remember?).

And finally, we can use that threshold to paint a map with the predicted range of the saguaro!
```{r, fig.dim= c(6,8)}
# Plot base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Only plot areas where probability of occurrence is greater than the threshold
plot(predict.presence.pseudo > bc.threshold, 
     add = TRUE, 
     legend = FALSE, 
     col = "olivedrab")

# And add those observations
points(x = obs.data$longitude, 
       y = obs.data$latitude, 
       col = "black",
       pch = "+", 
       cex = 0.75)

# Redraw those country borders
plot(great_britain_prj, add = TRUE, border = "grey5")
box()

```

Hmmm…that doesn’t look right. It plotted a large portion of the map green. Let’s look at what we actually asked R to plot, that is, we plot the value of predict.presence > bc.threshold. So what is that?

```{r}
predict.presence.pseudo > bc.threshold

```

The comparison of these two rasters produces another raster with values of only 0 or 1: 0 where the comparison evaluates as FALSE (i.e., when the value in a grid cell of predict.presence is less than or equal to the value in the corresponding grid cell of bc.threshold) and 1 where the comparison evaluates at TRUE. Since there are two values in this comparison (the 0 and 1 in the values field), we need to update what we pass to the col parameter in our plot call. Instead of just passing a single value, we provide a color for 0 (NA) and a color for 1 ("olivedrab"):

```{r, fig.dim= c(6,8)}
# Plot base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Only plot areas where probability of occurrence is greater than the threshold
plot(predict.presence.pseudo > bc.threshold, 
     add = TRUE, 
     legend = FALSE, 
     col = c(NA, "olivedrab"))

# And add those observations
points(x = obs.data$longitude, 
       y = obs.data$latitude, 
       col = "black",
       pch = "+", 
       cex = 0.75)

# Redraw those country borders
plot(great_britain_prj, add = TRUE, border = "grey5")
box()

```

A final note on our approach: the map we have drawn presents a categorical classification of whether a particular point on the landscape will be suitable or not for the species of interest. This classification relies quite heavily on the value of the threshold (see bc.threshold and the documentation for threshold) and the pseudo-absence points. Given that we used random sampling to generate those pseudo-absence points, there is potential for variation in the predicted range if you run this code more than once (try it! if you re-run the code from the point of creating the pseudo-absence points, you are almost guaranteed a different map.). There are a number of approaches to dealing with this variation, and the paper by Barbet-Massin et al. (2012) is a great resource. I’ll leave it as homework for you to determine which approach is most appropriate here!


#Advanced: Forecasting distributions
Now that you have a species distribution model, you can make predictions about the distribution under different climate scenarios. Let us pause for a moment and be very clear about this approach. With all kinds of math wizardry on our side, we are attempting to predict the future. Which means any predictions we make should be interpreted with extreme caution. If you are going to go about an approach such as this, it would be wise to run a variety of different models and a variety of different climate scenarios. There are links to such resources in the Additional Resources section, below.

#Forecast climate data
We will need to download climate data for the time period of interest. For the purposes of this lesson, we will look a climate projections for the year 2070. We will use the GFDL-ESM2G model with an RCP of 4.5 CO2. For an examination of different forecast models, see McSweeney et al. 2015. We can download the data like we did before, using raster’s getData() function, and send it do the data folder:


```{r}
#forecast.data <- getData(name = "CMIP5", # forecast data
                         #var = "bio",    # bioclim
                         #res = 5,      # 5 minute resolution
                         #path = "~/My R/Konservasi_ANJ/species_modelling", # destination directory
                         #model = "GD",   # 
                         #rcp = "45",     # CO2 increase 4.5
                         #year = 50)      # 2050

forecast.data <- cmip6_world("MIROC-ES2L", "370", "2041-2060","bioc", 5, "~/My R/Konservasi_ANJ/species_modelling/output/5_res")
```


We need to do one more thing before we can use our data, and that is to make sure the names in our model of bioclimatic variables line up with the names in the forecast data object

```{r}
forecast.data <- brick(forecast.data)
names(forecast.data) <- names(bioclim.data)

```

#Get out the crystal ball
Now that we have the forecast data, we can apply the model we build above, bc.model, to the forecast climate data:
```{r}
forecast.presence <- dismo::predict(bc.model, forecast.data, ext = geographic.extent)
```

If you want to look at the predicted probabilities of occurrence, you can modify the code we used above.

```{r, fig.dim= c(6,8)}
# Plot base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Add model probabilities
plot(forecast.presence, add = TRUE)

# Redraw those country borders
plot(great_britain_prj, add = TRUE, border = "grey5")

# Add original observations
points(obs.data$longitude, obs.data$latitude, col = "transparent", pch = 20, cex = 0.75)
box()

```

We can also map our predictions for presence / absence, using the same threshold that we did for predictions based on current climate data.
```{r, fig.dim= c(6,8)}
# Plot base map
plot(great_britain_prj, 
     xlim = c(min.lon, max.lon),
     ylim = c(50, 61),
     axes = TRUE, 
     col = "grey95")

# Only plot areas where probability of occurrence is greater than the threshold
plot(forecast.presence > bc.threshold, 
     add = TRUE, 
     legend = FALSE, 
     col = c(NA, "olivedrab"))

# And add those observations
#points(x = obs.data$longitude, 
    #   y = obs.data$latitude, 
     #  col = "black",
      # pch = "+", 
       #cex = 0.6)

# Redraw those country borders
plot(great_britain_prj, add = TRUE, border = "grey5")
box()

```

```{r}
print(bc.threshold)
```