---
title: "Non-Parametric Statistical Test"
subtitle: "Trembesi (Tinggi)"
author: "Robby"
date: "6/28/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This test is used when normality assumption of a dataset is not achieved.

```{r message=FALSE, warning=FALSE}
library(rcompanion)
library(multcompView)
library(DescTools)
library(FSA)
library(knitr)
```

### **Reading and preparing data**
```{r}
df <- read.delim("read data trembesi tinggi.txt")
kable(df)
```

```{r message=FALSE, warning=FALSE}
attach(df)
```

### **Kruskal-Wallis Rank Sum Test**
```{r}
kruskalw <- kruskal.test(Tinggi ~ Media)
print(kruskalw)
```

*Note: If Kruskal-Wallis test result is significant (p < 0.05) conitnue with post hoc test.

### **Post Hoc Tests**
*Test for several post hoc methods for comparison*

#### Pairwise Wilcoxon Rank Sum Tests
```{r error = T}
wilcox <- pairwise.wilcox.test(Tinggi,Media,
                               p.adjust.method = "BH", exact = FALSE)
#p.adjust.method: Returns p-values adjusted using Benjamini & Hochberg (1995) method ("BH")
wilcox1 <- wilcox$p.value
wilcox2 <- fullPTable(wilcox1)
wilcox3 <- multcompLetters(wilcox2,compare = "<", threshold = 0.05, 
                           Letters = letters,reversed = T)
print(list(wilcox_pairwise = wilcox, wilcox3))
```


#### Dunn's Kruskal-Wallis Multiple Comparisons.
**An error indicates that there's no significant differences**
```{r error=T}
Dunn <- dunnTest(Tinggi ~ Media, method = "bh")
Dunn1 <- Dunn$res
Dunn2 <- cldList(comparison = Dunn1$Comparison, p.value = Dunn1$P.adj, threshold = 0.05, reversed = T)
print(list(Dunn = Dunn, Dunn2))
```

#### Nemenyi Test
```{r}
Media <- as.factor(Media)
nemenyi <- NemenyiTest(x = Tinggi, g = Media, dist = "tukey")
print(nemenyi)
```

**An error from cldList() indicate that there's no significant differences**
```{r error= T}
try(cldList(comparison = nemenyi$Comparison, p.value = nemenyi$P.adj, threshold = 0.05, reversed = T))
```

