---
title: "Homogeneity and normality"
subtitle: "Trembesi (Tinggi)"
author: "Robby"
date: "6/22/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(FSA)
library(car)
library(nortest)
library(knitr)

```


### **Reading and preparing data**
```{r}
df <- read.delim("read data trembesi tinggi.txt")
kable(df)
dependent.var <- df[,2] #Column variabel dependent
independent.var <- df[,1]#Column variabel independent
satuan <- "(cm)"
```

### **Data descriptive summary**
```{r message=FALSE, warning=FALSE}
attach(df)
```

```{r}
df_summary <- data.frame(Summarize(Tinggi ~ Media, digits = 2))
kable(df_summary, format = "pipe")
```

### **Homogeneity test**

#### Levene's test for homogeneity of variance.
*if p > 0.05 assumption of homogeneity fulfilled*
```{r}
levene <- suppressWarnings(leveneTest(Tinggi ~ Media, data = df))
print(levene)
```

### **Normality test**
*if p > 0.05 assumption of normal distribution fulfilled*

#### Shapiro-Wilk Normality Test
```{r}
normalitas <- shapiro.test(Tinggi)
print(normalitas)
```

#### Kolmogorov-Smirnov Tests
```{r}
uji_ks <- suppressWarnings(ks.test(Tinggi,"pnorm",mean = mean(Tinggi), 
                            sd = sd(Tinggi)))
print(uji_ks)
```

#### Lilliefors (Kolmogorov-Smirnov) test for normality
Although the test statistic obtained from lillie.test(x) is the same as that obtained from ks.test(x, "pnorm", mean(x), sd(x)), it is not correct to use the p-value from the latter for the composite hypothesis of normality (mean and variance unknown), since the distribution of the test statistic is different when the parameters are estimated.

```{r}
lillie <- lillie.test(Tinggi)
print(lillie)
```

### **Q-Q Plot**
A Q-Q plot is a scatterplot created by plotting two sets of quantiles against one another. If both sets of quantiles came from the same distribution, we should see the points forming a line that’s roughly straight.
```{r, fig.width=5.8, fig.height=4.5}
kol2 <- colnames(df)[2]
QQplot <- qqPlot(Tinggi, id = F, ylab = kol2, main = "Quantile - Quantile Plot")
```

### **Box-and-whisker plot**
```{r fig.align='left', fig.width= 8, fig.height= 4}
box_whisker <- boxplot(Tinggi ~ Media, 
                        xlab = paste(kol2,satuan), ylab = "", col = "orange", border = 
                        "brown", horizontal = TRUE, las = 1, par(mar = c(5,12,1,1), font.axis=2, font.lab=2))
```


